### 2008-09-17 doc/client/client_technical.tex
Client documentation update

\documentclass{book}
%\documentclass{article}                            %for shorter notes
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Paper author},
  pdftitle = {Paper title},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\newcommand{\ACC}{\texttt{ACC}}
\newcommand{\TargetGenerator}{\texttt{TargetGenerator}}
\newcommand{\TargetRetriever}{\texttt{TargetRetriever}}
\newcommand{\ExecutionTarget}{\texttt{ExecutionTarget}}
\newcommand{\Broker}{\texttt{Broker}}
\newcommand{\Submitter}{\texttt{Submitter}}
\newcommand{\JobDescription}{\texttt{JobDescription}}
\newcommand{\JobSupervisor}{\texttt{JobSupervisor}}
\newcommand{\JobController}{\texttt{JobController}}
\newcommand{\Job}{\texttt{Job}}
\newcommand{\UserConfig}{\texttt{UserConfig}}
\newcommand{\ClientInterface}{\texttt{ClientInterface}}
\newcommand{\URL}{\texttt{URL}}
\newcommand{\XML}{\texttt{XML}}
\newcommand{\Loader}{\texttt{Loader}}
\newcommand{\ACCConfig}{\texttt{ACCConfig}}
\newcommand{\Logger}{\texttt{Logger}}
\newcommand{\XMLNode}{\texttt{XMLNode}}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

%%%%---- The title ----
{\centering \textsc{\Large ARC1lib, A Client Library for ARC1}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
{\centering \textit{\large Paper subtitle}\large \par}

@@

This document describes from a technical viewpoint a plugin based client library for the new 
Web Service (WS) based Advanced Resource Connector~\cite{arc} ARC middleware. The library consists of 
a set of C++ classes for

@@

All capabilities are enabled for three different grid flavours (Production ARC, ARC1 and gLite~\cite{glite}) 
through a modular design using plugins specialized for each supported middleware. Future 
extensions to support additional middlewares involve plugin compilation only i.e. no recompilation 
of main libraries or client.

Using the library a set of command line tools have been built which puts the library's functionality 
at the fingertips of users. While this documentation will illustrate how such command line tool can be 
built, the main documentation of the command line tools is given in the client user manual~\cite{ui}.

@@

\chapter{Functionality Overview}
\label{sec:FuncOver}
The new ARC1lib makes extensive use of plugins for command handling. These plugins are handled by a set 
of higher level classes which thus are the ones to offer the plugin functionality to external calls. In 
this section an overview of the library's main functionality is given which also introduces the most 
important higher level classes.
\section{Resource Discovery and Information Retrieval}
\label{sec:TargetDiscovery}
With the increasing number of grid clusters around the world, a reliable and fast resource discovery and 
information retrieval capability is of crucial importance for a user interface. The new ARC1lib resource 
discovery and information retrieval component consists of three classes; The {\TargetGenerator}, the 
{\TargetRetriever} and the {\ExecutionTarget}. Of these the {\TargetRetriever} is a base class for further grid 
middleware specific specialization (plugin).

@@

\begin{figure}[ht]
\centering{{\scalebox{0.75}{\includegraphics{TargetDiscovery.png}}}
\caption{\label{fig:ResDisc}Diagram depicting the resource discovery and information retrieval process} }
\end{figure}

\begin{enumerate}
\item{The {\TargetGenerator} takes three {\URL} arrays as input. First array contains individually selected computing services, while the 
second array contains individually selected index servers (1a). The last array contains a list of rejected services. This 
list is parsed through alias resolve before being inserted as filter to the storage of found services (1b).}
\item{The complete list of {\URL}s pointing to computing services and index servers.}
\item{For each {\URL} a {\TargetRetriever} plugin is loaded using the ARC loader. The {\TargetRetriever} is initialized 
with the {\URL} and the information about this {\URL} pointing to a computing service or an index server.}
\item{An external call is received calling for targets to be prepared. Call is propagated to all loaded {\TargetRetriever}s.}
\item{The call for targets is processed by each {\TargetRetriever} in parallel. If the {\TargetRetriever} is initialized with an 
index server {\URL} it first registers at the Index Servers store kept by the {\TargetGenerator}. If allowed to register, the index 
server is queried and the query result processed.}
\item{The TargetRetriver finds a registered service and contacts the Found Services store kept by the {\TargetGenerator} to check 
if the service in question has been found be another {\TargetRetriever}. (Services often registers at more than one index server, 
thus different {\TargetRetriever}s may discover the same service.)}
\item{If the {\TargetGenerator} allows registration in the Found Services store, the {\TargetRetriever} progresses to interrogate the service.}
\item{Found and processed {\ExecutionTarget}s are stored in the Found Targets store kept by the {\TargetGenerator} for later usage 
(e.g. status printing or job submission).}
\item{Index servers often registers with other index servers creating a nested hierarchical information structure. For this 
reason a {\TargetRetriever} processing the query result from an index server may encounter other index servers. In such
events the {\TargetRetriever} will call itself recursively creating a new thread for the new {\URL}. This ensures optimum speed and 
performance for resource discovery.}
\end{enumerate}

The example above outlines how the {\TargetRetriever} works when initialized with an {\URL} pointing to and index server. However, 
as indicated it also accepts initialization with a service {\URL}s, in which case the {\TargetRetriever} will try to register the 
service in the {\TargetGenerator} Found Services store. If allowed the {\TargetRetriever} will progress with the target interrogation 
method.

\section{Job Submission}
\label{sec:JobSubmission}
Job submission starts with the resource discovery and target preparation as outlined in the Section~\ref{sec:TargetDiscovery}. 
Not until a list of possible targets (which allows the user) is available is the job description read in order to enable bulk 
job submission of widely different jobs without having to reperform the resource discovery. In addition to the classes mentioned 
above the job submission makes use of the {\Broker}, {\JobDescription} and {\Submitter} classes. At the time of writing the {\Broker} is 
not implemented, but it's functionality and interface is well known, and thus we will include it in the overview given here.
The {\Submitter} is base class for further grid middleware specific specialization (plugin) similarly to the {\TargetRetriever}.

@@

\section{Job Management}
Once a job is submitted job related information (job identification string, cluster etc) is stored in a local xml file which hosts 
similar information for all active jobs. This file may contain jobs running on completely different grid flavour, and thus job 
management should be handled using plugins similar to resource discovery and job submission. The job managing plugin is called 
the {\JobController} and it is supported by the {\JobSupervisor} and {\Job} classes.

@@

\begin{enumerate}
\item{The {\JobSupervisor} takes three arguments Jobs (job identification strings), ClusterSelect and ClusterReject.}
\item{A job identification string does not uniquely define which grid flavour runs a certain job. Thus this information is stored 
in joblist.xml upon submission by the {\Submitter} and joblist.xml is extensively used by the {\JobSupervisor} to identify the 
{\JobController} flavours which are to be loaded.}
\item{Suitable {\JobController}s are loaded}
\item{Jobs, ClusterSelect and ClusterReject are passed to the {\JobController} which uses the information to 
fill its internal JobStore.}
\item{Residing within the {\JobSupervisor} the {\JobController}s are now accessible for external calls (i.e. job handling).}
\end{enumerate}

\chapter{Implementation}
\label{sec:Implementation}
Clever text here.
\section{Generic Classes}
\subsection{{\ACC}}
The Arc Client Component ({\ACC}) class is a completely empty base class needed for the {\Loader} 
in order to create loadable classes.

\begin{shaded}
\begin{verbatim}
class ACC {
protected:
  ACC() {}
public:
  virtual ~ACC() {}
};
\end{verbatim}
\end{shaded}

\subsection{{\TargetGenerator}} The {\TargetGenerator} class is the main class for resource discovery and information retrieval. It loads 
{\TargetRetriever} plugins in accordance with the input {\URL}s using the ARC {\Loader}~\cite{hed} (e.g. if an {\URL} pointing to an ARC1 
resource is given the TargetRetrieverARC1 is loaded).

@@

\begin{shaded}
\begin{verbatim}
TargetGenerator(const std::list<std::string>& clusterselect,
                const std::list<std::string>& clusterreject,
                const std::list<std::string>& indexurls);
\end{verbatim}
\end{shaded}

where \texttt{clusterselect}, \texttt{clusterreject} and \texttt{indexurls} are lists of strings where each string is either 
an {\URL} or a regular string defined as an alias in the user (or system) configuration. Without any further explanation it is 
hard to decode from an {\URL} if it points to a index server or computing service of a certain grid flavour. Thus in order for 
the {\TargetGenerator} to know which {\TargetRetriever} should be loaded {\URL}s have to be given in a format where the grid 
flavour precedes the actual {\URL}, e.g:

\begin{shaded}
\begin{small}
\begin{verbatim}
ARC0:ldap://grid.tsl.uu.se:2135/mds-vo-name=sweden,O=grid
ARC0:ldap://grid.tsl.uu.se:2135/nordugrid-cluster-name=grid.tsl.uu.se,Mds-Vo-name=local,o=grid
\end{verbatim}
\end{small}
\end{shaded}

where the former is the {\URL} to an index server while the latter is an {\URL} to a computing service.

@@

The {\TargetGenerator} will pass this request to the loaded {\TargetRetriever}s running them as individual threads 
for improved performance. The {\TargetGenerator} keeps records of by {\TargetRetriever} found index servers and computing 
services in order to avoid multiple identical queries. Accepted {\ExecutionTarget}s are stored in the FoundTargets 
array kept by the {\TargetGenerator}. See also Section~\ref{sec:TargetDiscovery} for a schematic drawing of the 
resource discovery process.

@@

\subsection{{\TargetRetriever}} The {\TargetRetriever} is base class for {\TargetRetriever} grid middleware specializations and 
inherits from the \texttt{ACC} base class in order to be loadable by the ARC \texttt{Loader}. It is designed to work on conjunction 
with the {\TargetGenerator} contains and contains the pure virtual method

@@

which is to be implemented by the specialized class. While it is not mandatory it is recommended that the specialized class 
divides this method into two components: QueryIndex and InterrogateTarget. The former is handles index server queries, and the 
latter the computing service queries and {\ExecutionTarget} preparation.

If an index server query yields an url to a different index server then the one queried, then the {\TargetRetriever} should 
call itself recursively creating a new {\TargetRetriever} for the discovered index server.

\subsection{{\ExecutionTarget}} The {\ExecutionTarget} is the class representation of a computing resource (queue) capable of 
executing a grid job. It serves as input to the {\Broker} which is foreseen to be able to select between different 
{\ExecutionTarget}s from different grid flavours without apriori knowing their difference. The {\ExecutionTarget} class 
mimics the glue2 information model (with a flattened structure), and thus a mapping between attributes from other information 
systems into the glue2 format is needed.  Appendix XX shows the current mapping for the production ARC, ARC1 and gLite middlewares.

@@

\begin{shaded}
\begin{verbatim}
Submitter *GetSubmitter() const;
\end{verbatim}
\end{shaded}

\subsection{{\Broker}} The {\Broker} is not yet implemented and thus not described here.

\subsection{{\JobDescription}} When addressing interoperability it is of paramount importance to transparently address grid job 
descriptions written in different job description languages by translating them automatically. In the ARC1lib library this functionality 
is implemented in the {\JobDescription} class. This is a generic class that takes a job description (string) as input

\begin{shaded}
\begin{verbatim}
setSource( const std::string source ) throw(JobDescriptionError);
\end{verbatim}
\end{shaded}

in any supported formats (currently xRSL, JDL, JSDL), converts and stores it in a JSDL-like internal job description format. The identification 
of the source's description format is internally handled by a \texttt{JobDescriptionOrderer} class which identifies the format by pattern
matching.

@@

\begin{shaded}
\begin{verbatim}
getProduct(std::string& product, std::string format="JSDL") throw(JobDescriptionError);
\end{verbatim}
\end{shaded}

@@

\subsection{{\Submitter}} {\Submitter} is base class for grid specific specializations (plugin). It submits job(s) to the service 
it represents and uploads (by the job needed) local input files.

@@

The \texttt{Submit} method fills the {\XML} node \texttt{info} with all needed information about the job for later job management. 
The {\Submitter} is returned by the {\ExecutionTarget} selected for job execution and thus the {\ExecutionTarget} populates 
(through {\XML} config element) the {\Submitter} with information about submission endpoint ({\URL}) and job description languages 
``spoken'' by the target.

\subsection{{\JobSupervisor}} The {\JobSupervisor} is responsible for loading the appropriate {\JobController}s for managing 
jobs running on a certain grid flavour. Job manipulation can be performed either on individual jobs or on groups of jobs (e.g. 
all jobs running on certain cluster or all jobs with job state ``FINISHED''), and in order to translate the by user given 
information into a set of loadable {\JobController}s the {\JobSupervisor} makes extensive use of the local {\XML} file housing 
information about all active jobs. Thus the {\JobSupervisor} constructor becomes

\begin{shaded}
\begin{verbatim}
JobSupervisor(const std::list<std::string>& jobs,
              const std::list<std::string>& clusterselect,
              const std::list<std::string>& clusterreject,
              const std::string joblist);
\end{verbatim}
\end{shaded}

where \texttt{jobs} is a list of job identification strings (jobids), \texttt{clusterselect} is a list of selected computing services, 
\texttt{clusterreject} is a list of rejected computing services and \texttt{joblist} is a string pointing to the local {\XML} file 
with job information. Syntax for \texttt{clusterselect} and \texttt{clusterreject} is identical to the one used by the {\TargetGenerator}
for for internal consistency.

@@

\begin{shaded}
\begin{verbatim}
std::list<Arc::JobController*> GetJobControllers(){return JobControllers;}
\end{verbatim}
\end{shaded}

@@

\subsection{{\JobController}} The {\JobController} is both base class for grid specific specializations, but also the implementor of all 
public functionality offered by the {\JobController}s. In other words all virtual functions of the {\JobController} are private. The 
initialization of a (specialized) {\JobController} object takes two steps. First the constructor has to invoked

\begin{shaded}
\begin{verbatim}
JobController(Arc::Config *cfg, std::string flavour);
\end{verbatim}
\end{shaded}

This is usually done by the ARC {\Loader} which sees to that the {\JobController} receives information about its flavour (grid) and 
the local {\XML} file containing information about all active jobs (flavour independent). Next step is the filling of the {\JobController}'s
job pool \texttt{JobStore} which is the pool of jobs that the {\JobController} can manage.

\begin{shaded}
\begin{verbatim}
void FillJobStore(std::list<std::string> jobs,
                  std::list<std::string> clusterselect,
                  std::list<std::string> cluterreject);
\end{verbatim}
\end{shaded}

Here \texttt{jobs}, \texttt{clusterselect} and \texttt{clusterreject} have the same interpretation as in the {\JobSupervisor} constructor. 
However, while the latter only had to identify which {\JobController}s should be loaded, the {\JobController} itself has to decode the 
supplied lists into a list of jobs with non-overlapping entries which will be filled into the JobStore. Due to possible conflicts 
within the supplied lists, rules have to be set up for dealing with contradictions:

\begin{enumerate}
\item{If the \texttt{jobs} list has entries, fill \texttt{JobStore} with the by user requested jobs.}
\item{If the \texttt{clusterselect} list has entries, fill \texttt{JobStore} with the jobs running on the selected clusters.}
\item{If clusters are rejected and \texttt{JobStore} has entries, remove from \texttt{JobStore} all jobs running on rejected clusters.}
\item{If \texttt{jobs} and \texttt{clusterselect} are both empty lists, fill \texttt{JobStore} with all jobs except those 
running on possible rejected clusters.}
\end{enumerate}

@@

\begin{shaded}
\begin{small}
\begin{verbatim}
void JobController::Stat(const std::list<std::string>& jobs,
                    const std::list<std::string>& clusterselect,
                    const std::list<std::string>& clusterreject,
                    const std::list<std::string>& status,
                    const bool longlist,
                    const int timeout){

>>>>

\documentclass{book}
%\documentclass{article}           %for shorter notes
\usepackage{graphicx}              %for PNG images (pdflatex)
%\usepackage{graphics}              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref}  %for \url tag
\usepackage{color}                 %for defining custom colors
\usepackage{framed}                %for shaded and framed paragraphs
\usepackage{textcomp}              %for various symbols, e.g. Registered Mark
\usepackage{geometry}              %for defining page size
\usepackage{longtable}             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Paper author},
  pdftitle = {Paper title},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\newcommand{\arclib}{libarcclient}
\newcommand{\ACC}{\texttt{ACC}}
\newcommand{\ACCConfig}{\texttt{ACCConfig}}
\newcommand{\Broker}{\texttt{Broker}}
\newcommand{\ClientInterface}{\texttt{ClientInterface}}
\newcommand{\ClientTCP}{\texttt{ClientTCP}}
\newcommand{\ClientHTTP}{\texttt{ClientHTTP}}
\newcommand{\ClientSOAP}{\texttt{ClientSOAP}}
\newcommand{\ExecutionTarget}{\texttt{ExecutionTarget}}
\newcommand{\Job}{\texttt{Job}}
\newcommand{\JobController}{\texttt{JobController}}
\newcommand{\JobDescription}{\texttt{JobDescription}}
\newcommand{\JobSupervisor}{\texttt{JobSupervisor}}
\newcommand{\Loader}{\texttt{Loader}}
\newcommand{\Logger}{\texttt{Logger}}
\newcommand{\TargetGenerator}{\texttt{TargetGenerator}}
\newcommand{\TargetRetriever}{\texttt{TargetRetriever}}
\newcommand{\Submitter}{\texttt{Submitter}}
\newcommand{\URL}{\texttt{URL}}
\newcommand{\UserConfig}{\texttt{UserConfig}}
\newcommand{\XML}{\texttt{XML}}
\newcommand{\XMLNode}{\texttt{XMLNode}}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

%%%%---- The title ----
{\centering \textsc{\Large {\arclib}}\Large \par}
\vspace*{0.5cm}

%%%%---- A subtitle, if necessary ----
{\centering \textit{\large A Client Library for ARC}\large \par}

@@

This document describes from a technical viewpoint a plugin based
client library for the new Web Service (WS) based Advanced Resource
Connector~\cite{arc} (ARC) middleware. The library consists of a set
of C++ classes for

@@

All capabilities are enabled for three different grid flavours
(Production ARC, ARC1 and gLite~\cite{glite}) through a modular design
using plugins specialized for each supported middleware. Future
extensions to support additional middlewares involve plugin
compilation only i.e.\ no recompilation of main libraries or clients
is necessary.

Using the library, a set of command line tools have been built which
puts the library's functionality at the fingertips of users. While
this documentation will illustrate how such command line tools can be
built, the main documentation of the command line tools is given in
the client user manual~\cite{ui}.

@@

\chapter{Functionality Overview}
\label{sec:FuncOver}

The new {\arclib} makes extensive use of plugins for command
handling. These plugins are handled by a set of higher level classes
which thus are the ones to offer the plugin functionality to external
calls. In this section an overview of the library's main functionality
is given which also introduces the most important higher level
classes.

\section{Resource Discovery and Information Retrieval}
\label{sec:TargetDiscovery}

With the increasing number of grid clusters around the world, a
reliable and fast resource discovery and information retrieval
capability is of crucial importance for a user interface. The new
{\arclib} resource discovery and information retrieval component
consists of three classes; the {\TargetGenerator}, the
{\TargetRetriever} and the {\ExecutionTarget}. Of these the
{\TargetRetriever} is a base class for further grid middleware
specific specialization (plugin).

@@

\begin{figure}[ht]
\centering{{\scalebox{0.75}{\includegraphics{TargetDiscovery.png}}}
\caption{\label{fig:ResDisc}Diagram depicting the resource discovery
  and information retrieval process}}
\end{figure}

\begin{enumerate}
\item{The {\TargetGenerator} takes three arguments as input. The first
  argument is a reference to a {\UserConfig} object containing a
  representing the contents of the user's configuration file. The
  second and third arguments contain lists of strings. The first list
  contains individually selected and rejected computing services,
  while the second list contains individually selected and rejected
  index servers. Rejected servers and services are identified by that
  its name is prefixed by a minus sign in the lists. The name of the
  servers and services should be given either in the form of an alias
  defined in the {\UserConfig} object or as the name of its grid
  flavour followed by a colon and the URL of its information contact
  endpoint.}
\item{These lists are parsed through alias resolution before being
  used to initialize the complete list of selected and rejected
  {\URL}s pointing to computing services and index servers.}
\item{An external call is received calling for targets to be
  prepared.}
\item{For each selected index server and computing service a
  {\TargetRetriever} plugin for the server's or service's grid flavour
  is loaded using the ARC loader. The {\TargetRetriever} is
  initialized with its {\URL} and the information about whether it
  represents a computing service or an index server. The call for
  targets is processed by each {\TargetRetriever} in parallel.}
\item{A {\TargetRetriever} representing an index server first tries to
  register at the index server store kept by the {\TargetGenerator}.
  If allowed to register, the index server is queried and the query
  result processed. The {\TargetGenerator} will not allow
  registrations from index servers present in its list of rejected
  index servers or from servers that have already registered
  once. Index servers often registers at more than one index server,
  thus different {\TargetRetriever} may discover the same server.)}
\item{If while processing the query result the {\TargetRetriever}
  finds an other registered index server or a registered computing
  service it creates an a new {\TargetRetriever} for the found server
  or service and forwards the call for targets to the new
  {\TargetRetriever}.}
\item{A {\TargetGenerator} representing a computing service first
  tries to register at the service store kept by the
  {\TargetGenerator}. If allowed to register, the computing server is
  queried and the query result processed. The {\TargetGenerator} will
  not allow registrations from computing services present in its list
  of rejected computing services or from service that have already
  registered once. Computing and services often registers at more than
  one index server, thus different {\TargetRetriever} may discover the
  same service.)}
\item{When processing the query result the {\TargetRetriever} will
  create an {\ExecutionTarget} for each queue found on the computing
  service and collect all possible information about them. It will
  then store the {\ExecutionTarget} in the found targets store kept
  by the {\TargetGenerator} for later usage (e.g. status printing or
  job submission).}
\end{enumerate}

\section{Job Submission}
\label{sec:JobSubmission}

Job submission starts with the resource discovery and target
preparation as outlined in the Section~\ref{sec:TargetDiscovery}.  Not
until a list of possible targets (which allows the user) is available
is the job description read in order to enable bulk job submission of
widely different jobs without having to reperform the resource
discovery. In addition to the classes mentioned above the job
submission makes use of the {\Broker}, {\JobDescription} and
{\Submitter} classes. At the time of writing the {\Broker} is not
implemented, but it's functionality and interface is well known, and
thus we will include it in the overview given here.  The {\Submitter}
is base class for further grid middleware specific specialization
(plugin) similarly to the {\TargetRetriever}.

@@

\section{Job Management}

Once a job is submitted job related information (job identification
string, cluster etc) is stored in a local xml file which hosts similar
information for all active jobs. This file may contain jobs running on
completely different grid flavours, and thus job management should be
handled using plugins similar to resource discovery and job
submission. The job managing plugin is called the {\JobController} and
it is supported by the {\JobSupervisor} and {\Job} classes.

@@

\begin{enumerate}
\item{The {\JobSupervisor} takes four arguments as input. The first
  argument is a reference to a {\UserConfig} object containing a
  representing the contents of the user's configuration file. The
  second is a list of strings containing job identifiers and job
  names, the third is a list of strings of clusters to select or
  reject (in the same format as described for the {\TargetGenerator}
  above). The last argument is the name of the file containing the
  local information about active jobs, hereafter called the joblist
  file.}
\item{A job identifier does not uniquely define which grid flavour
  runs a certain job. Thus this information is stored in the joblist
  file upon submission by the {\Submitter} and the joblist file is
  extensively used by the {\JobSupervisor} to identify the
  {\JobController} flavours which are to be loaded. The information in
  the joblist file is also used to lookup the job identifier for jobs
  specified using job names. Alias resolution for the selected and
  rejected clusters are performed using the information in the
  {\UserConfig} object.}
\item{Suitable {\JobController}s are loaded}
\item{The list of job identifiers and selected and rejected clusters
  are passed to each {\JobController} which uses the information to
  fill its internal JobStore.}
\item{Residing within the {\JobSupervisor} the {\JobController}s are
  now accessible for external calls (i.e. job handling).}
\end{enumerate}

\chapter{Implementation}
\label{sec:Implementation}

In this section an overview of all important classes in the new
{\arclib} is given. The overview is subdivided in two parts where
first all the generic classes are presented,
Section~\ref{sec:genclass}, before the grid flavour specializations
are presented in Section~\ref{sec:plugins}. Classes are described both
in words and by code.

\section{Generic Classes}
\label{sec:genclass}

\subsection{{\ACC}}

The Arc Client Component ({\ACC}) class is a base class needed for the
{\Loader} in order to create loadable classes. It stored information
about which flavour it supports and security related information
needed by all ACC specializations.

\begin{shaded}
\begin{verbatim}
class ACC {
protected:
  ACC(Config *cfg, const std::string& flavour);
public:
  virtual ~ACC();
  const std::string& Flavour();
protected:
  std::string flavour;
  std::string proxyPath;
  std::string certificatePath;
  std::string keyPath;
  std::string caCertificatesDir;
};
\end{verbatim}
\end{shaded}

\subsection{{\TargetGenerator}}

 The {\TargetGenerator} class is the main class for resource discovery
 and information retrieval. It loads {\TargetRetriever} plugins in
 accordance with the input {\URL}s using the ARC {\Loader}~\cite{hed}
 (e.g. if an {\URL} pointing to an ARC1 resource is given the
 TargetRetrieverARC1 is loaded).

@@

\begin{shaded}
\begin{verbatim}
    TargetGenerator(const UserConfig& usercfg,
		    const std::list<std::string>& clusters,
		    const std::list<std::string>& indexurls);
\end{verbatim}
\end{shaded}

where \texttt{clusters}, \texttt{indexurls} are lists of strings where
each string is either the name of a grid flavour followed by a colon
and the URL of an information contact endpoint for a computing service
or index server, respectively, of that grid flavour or an alias
defined in the {\UserConfig} object passed as the first
argument, e.g:

\begin{shaded}
\begin{small}
\begin{verbatim}
ARC0:ldap://grid.tsl.uu.se:2135/Mds-Vo-name=Sweden,o=grid
ARC0:ldap://grid.tsl.uu.se:2135/nordugrid-cluster-name=grid.tsl.uu.se,Mds-Vo-name=local,o=grid
\end{verbatim}
\end{small}
\end{shaded}

where the former is the URL to an index server while the latter is an
URL to a computing service.

@@

The {\TargetGenerator} will pass this request to the loaded
{\TargetRetriever}s running them as individual threads for improved
performance. The {\TargetGenerator} keeps records of index servers and
computing services found by the {\TargetRetriever}s in order to avoid
multiple identical queries. Accepted {\ExecutionTarget}s are stored in
the FoundTargets array kept by the {\TargetGenerator}. See also
Section~\ref{sec:TargetDiscovery} for a schematic drawing of the
resource discovery process.

@@

\subsection{{\TargetRetriever}}

The {\TargetRetriever} is base class for {\TargetRetriever} grid
middleware specializations and inherits from the \texttt{ACC} base
class in order to be loadable by the ARC \texttt{Loader}. It is
designed to work in conjunction with the {\TargetGenerator} and
contains the pure virtual method

@@

which is to be implemented by the specialized class. While it is not
mandatory it is recommended that the specialized class divides this
method into two components: QueryIndex and InterrogateTarget. The
former handles index server queries, and the latter the computing
service queries and {\ExecutionTarget} preparation.

If an index server query yields an url to a different index server
than the one queried, then the {\TargetRetriever} should call itself
recursively creating a new {\TargetRetriever} for the discovered index
server.

\subsection{{\ExecutionTarget}}

The {\ExecutionTarget} is the class representation of a computing
resource (queue) capable of executing a grid job. It serves as input
to the {\Broker} which is foreseen to be able to select between
different {\ExecutionTarget}s from different grid flavours without a
priori knowing their difference. The {\ExecutionTarget} class mimics
the glue2 information model (with a flattened structure), and thus a
mapping between attributes from other information systems into the
glue2 format is needed. Appendix XX shows the current mapping for the
production ARC, ARC1 and gLite middlewares.

@@

\begin{shaded}
\begin{verbatim}
Submitter *GetSubmitter(const UserConfig& ucfg) const;
\end{verbatim}
\end{shaded}

The {\UserConfig} object is used to configure the security related
information needed by the {\Submitter}.

\subsection{{\Broker}}

The {\Broker} is not yet implemented and thus not described here.

\subsection{{\JobDescription}}

When addressing interoperability it is of paramount importance to
transparently address grid job descriptions written in different job
description languages by translating them automatically. In the
{\arclib} library this functionality is implemented in the
{\JobDescription} class. This is a generic class that takes a job
description (string) as input

\begin{shaded}
\begin{verbatim}
void setSource(const std::string source);
\end{verbatim}
\end{shaded}

in any supported formats (currently xRSL~\cite{xrsl}, JDL~\cite{jdl},
JSDL~\cite{jsdl}), converts and stores it in a JSDL-like internal job
description format. The identification of the source's description
format is internally handled by a \texttt{JobDescriptionOrderer} class
which identifies the format by pattern matching.

@@

\begin{shaded}
\begin{verbatim}
void getProduct(std::string& product, std::string format = "JSDL")
\end{verbatim}
\end{shaded}

@@

\subsection{{\Submitter}}

{\Submitter} is base class for grid specific specializations (plugin).
It submits job(s) to the service it represents and uploads (by the job
needed) local input files.

@@

The \texttt{Submit} method fills the {\XMLNode} \texttt{info} with all
needed information about the job for later job management. The
{\Submitter} is returned by the {\ExecutionTarget} selected for job
execution and thus the {\ExecutionTarget} populates (through its
{\XMLNode} config element) the {\Submitter} with information about
submission endpoint ({\URL}) and job description languages understood
by the target.

\subsection{{\JobSupervisor}}

The {\JobSupervisor} is responsible for loading the appropriate
{\JobController}s for managing jobs running on a certain grid
flavour. Job manipulation can be performed either on individual jobs
or on groups of jobs (e.g.\ all jobs running on certain cluster or all
jobs with job state ``FINISHED''), and in order to translate the
information given by the user into a set of loadable {\JobController}s
the {\JobSupervisor} makes extensive use of the local joblist file
housing information about all active jobs. Thus the {\JobSupervisor}
constructor becomes

\begin{shaded}
\begin{verbatim}
JobSupervisor(const UserConfig& usercfg,
              const std::list<std::string>& jobs,
              const std::list<std::string>& clusters,
              const std::string& joblist);
\end{verbatim}
\end{shaded}

where \texttt{jobs} is a list of job identifiers and job names,
\texttt{clusters} is a list of selected and rejected computing
services in the same format as described above for the
{\TargetGenerator}, \texttt{joblist} is a string containing the name
of the joblist file.

@@

\begin{shaded}
\begin{verbatim}
const std::list<Arc::JobController*>& GetJobControllers();
\end{verbatim}
\end{shaded}

@@

\subsection{{\JobController}}

The {\JobController} is a base class for grid specific
specializations, but also the implementor of all public functionality
offered by the {\JobController}s. In other words all virtual functions
of the {\JobController} are private. The initialization of a
(specialized) {\JobController} object takes two steps. First the
{\JobController} specialization for the required grid flavour must be
loaded by the ARC {\Loader}, which sees to that the {\JobController}
receives information about its flavour (grid) and the local joblist
file containing information about all active jobs (flavour
independent). Next step is the filling of the {\JobController}'s job
pool \texttt{JobStore} which is the pool of jobs that the
{\JobController} can manage.

\begin{shaded}
\begin{verbatim}
void FillJobStore(const std::list<URL>& jobids,
                  const std::list<URL>& clusterselect,
                  const std::list<URL>& cluterreject);
\end{verbatim}
\end{shaded}

Here \texttt{jobids}, \texttt{clusterselect} and
\texttt{clusterreject} have been resolved for job names and aliases by
the {\JobSupervisor}, and no further resolution is needed. The
following rules are observed when filling the \texttt{JobStore}:

\begin{enumerate}
\item{If the \texttt{jobids} list has entries, fill \texttt{JobStore}
  with the by user requested jobs.}
\item{If the \texttt{clusterselect} list has entries, fill
  \texttt{JobStore} with the jobs running on the selected clusters.}
\item{If clusters are rejected and \texttt{JobStore} has entries,
  remove from \texttt{JobStore} all jobs running on rejected
  clusters.}
\item{If \texttt{jobs} and \texttt{clusterselect} are both empty
  lists, fill \texttt{JobStore} with all jobs except those running on
  possible rejected clusters.}
\end{enumerate}

@@

\begin{shaded}
\begin{small}
\begin{verbatim}
bool JobController::Stat(const std::list<std::string>& status,
                         const bool longlist,
                         const int timeout) {
